A Survey of Current Exoskeletons\\ and Their Control Architectures and Algorithms\\ (Draft 3.2)

### 2015-09-30 survey.tex

\begin{abstract}
In the simplest case, where the environment around a robotic system remains relatively fixed, controlling the dynamic interaction between the robot and the environment to achieve a specific physical objective is difficult.  In the case where an operator effectively wears the robot, i.e., a robotic exoskeleton, and would like to move through unknown and dynamically evolving environments, the situation becomes very complicated; the robotic system needs to account for the dynamics of multiple fundamentally linked systems that are for the most part difficult to model.  The goal of this document is to provide a survey of the literature on the design, modeling, and control of robotic exoskeletons and to highlight what we believe are several of the most promising currently existing approaches to the complicated goal of enabling dynamic yet safe interaction between operators, exoskeletons, and the environment.  Where information was available, we summarize specific contributions related to actuation, power storage and generation, sensor type and distribution, mechanical architecture, as well as control system design for a number of different exoskeleton hardware systems.  Included as part of this report are recommendations on what we see as helpful components as well as potentially problematic issues in current systems and control approaches. 
\end{abstract}

@@

We then list other exoskeletons in a variety of categories.

@@

Sensitivity amplification strategies circumvent many of the direct operator-robot sensing issues by estimating human intent from measurements of the exoskeleton hardware. The issue with this policy is that it cannot generally distinguish between user generated torques and those produced by external sources.  As such, they end up amplifying not only the user's intent, but also external disturbances, requiring the operator to expend valuable energy to stabilize the system.  Additionally, even under the assumption that external forces can be minimized or mitigated by the user, enabling the suit to be highly responsive to perceived user intent implies that the hardware needs to be operating near the margin of stability.  This is not an optimal strategy for highly dynamic behaviors, as the suit may be pushed beyond the point to which the operator can reasonably stabilize it.

In light of these opinions, our main recommendation is that a control policy which implements a \emph{modified get out of the way} strategy would provide a reasonable strategy for dynamic exoskeleton control.  The get out of the way controller used in the Hydraulic Lower Extremity Exoskeleton discussed above would provide a starting point for this strategy.  This controller utilizes IMUs, foot contact sensors, and joint torque sensors to implement a user-reactive gravity compensating strategy.  Augmenting the sensor suit and control system to include sensors that directly measure the forces between the operator and the robot would enable this system not only avoid the user, but also to reject external disturbances.

To achieve this type of augmented get out of the way control, it may be possible to use low information density EMG (thresholded on/off muscle activation signals) with robot force/torque and angular measurements to filter operator-robot interaction forces.  As a longer term solution, we highlight research in the area of MEMs, nano-fabrication, soft sensors, etc., which has lead to the development of new, flexible ``robotic skin'' sensing arrays.  Integrating these sensing arrays into operator clothing (or attaching directly to the exoskeleton), it may be possible to directly measure human interaction forces (i.e., estimating pressure / force and exoskeleton contact locations), which was not feasible in the recent past.  Dense sensing arrays add redundancy and can improve measurement reliability.

We also highly stress the importance of high-bandwidth control loops.  It is important that an exoskeleton's control bandwidth stays well above (typically $2-3\times$) the human control bandwidth ($2-5\,$Hz), to track dynamic motion.  Several of exoskeletons controllers are close to lower limit of this human bandwidth threshold under no-load conditions, and so can interfere in dynamic tasks.  Note that these bandwidth issues are primarily a function of the exoskeleton hardware, and thus special care/analysis needs to be given/conducted to make sure that the hardware is capable of physically producing required behaviors.

Lastly, the importance of producing hardware that is physically capable of achieving dynamic performance is fundamentally linked to how mass is distributed throughout the system.  In general, this means that minimizing mass, more specifically distal masses on extremities, is desirable, which in turn means that at certain degrees of freedom must remain unactuated.  Underactuted control is a complicated problem which can either be ignored, in the sense that the control algorithm assumes the user will provide the necessary forces and torques for the unactuated degrees of freedom, dealt with using optimization-based control methods, or can be partially compensated using passive energy storage and dissipation elements.  Our recommendation is that a combination of biomechanically-inspired passive elements and optimization-based control methods will ultimately produce the best strategy for dealing with underactuation in higly-dynamic exoskeleton systems.

@@

% Finally, we note that 
%Many of the common model-based control methods in this survey employ feedback linearization and inverse dynamics strategies for dynamic exoskeleton control.  
%These methods (incorrectly) assume the exoskeleton is fully actuated and require users to provide certain torques at uncontrolled joints.  There are more advanced control schemes that extend to underactuated systems and may apply here (e.g., optimization-based controllers or virtual model control).  However, we note that there is a balance between feedback and feedforward control.  The better one is, the less reliant the system is on the other.
%
%Improved feedback and direct sensing of user intent / human interaction forces therefore also offers to reduce sensitivity to high accuracy models and the fully actuated assumptions in inverse dynamics control.

>>>>

\begin{abstract}
In the simplest case, where the environment around a robotic system remains relatively fixed, controlling the dynamic interaction between the robot and the environment to achieve a specific physical objective is difficult.  In the case where an operator effectively wears the robot, i.e., a robotic exoskeleton, and would like to move through unknown and dynamically evolving environments, the situation becomes very complicated; the robotic system needs to account for the dynamics of multiple fundamentally linked systems that are, for the most part, difficult to model.  The goal of this document is to provide a survey of the literature on the design, modeling, and control of robotic exoskeletons and to highlight what we believe are several of the most promising currently existing approaches to the complicated goal of enabling dynamic yet safe interaction between operators, exoskeletons, and the environment.  Where information was available, we summarize specific contributions related to actuation, power storage and generation, sensor type and distribution, mechanical architecture, as well as control system design for a number of different exoskeleton hardware systems.  Included as part of this report are recommendations on what we see as helpful components as well as potentially problematic issues in current systems and control approaches. 
\end{abstract}

@@

\noindent Exoskeletons in a variety of categories are listed in the Appendix.\\

\noindent Based on our review, we have made the following conclusions:
\begin{itemize}
\item Most exoskeletons estimate user intent indirectly, from measurements of suit variables.
\item Due to limitations in current sensory systems, it is difficult to sense user intent directly. 
\item As direct measurement of user intent is critical in dynamic, noisy environments, we recommend better fusion of low information density sensory data obtained from users with closed-loop control based on suit variables.
\item In the longer term, we recommend investigation and development of more advanced human-exoskeleton sensing technology to capture user intent.
\end{itemize}

@@

Sensitivity amplification strategies circumvent many of the direct operator-robot sensing issues by estimating human intent from measurements of the exoskeleton hardware. The issue with this policy is that it cannot generally distinguish between user generated torques and those produced by external sources.  As such, they end up amplifying not only the user's intent, but also external disturbances, requiring the operator to expend valuable energy to stabilize the system.  
In dynamic situations the operator may not be capable of actively stabilizing the exoskeleton.
% Additionally, even under the assumption that external forces can be minimized or mitigated by the user, enabling the suit to be highly responsive to perceived user intent implies that the hardware needs to be operating near the margin of stability.  This is not an optimal strategy for highly dynamic behaviors, as the suit may be pushed beyond the point to which the operator can reasonably stabilize it.

% Based on these observations, our main recommendation is that a \emph{modified get out of the way} control strategy provides a reasonable policy for dynamic exoskeletons.
Based on these observations, we recommend (at least in the short term) fusing the two predominant approaches to exoskeleton control. That is, combining sensitivity amplification strategies with direct user measurement to robustly distinguish user intent from external disturbances.
% For instance, the get out of the way controller used in the Hydraulic Lower Extremity Exoskeleton would provide a starting point for this strategy.  This controller utilizes IMUs, foot contact sensors, and joint torque sensors to implement a user-reactive gravity compensating strategy that does not directly measure the user.  Augmenting the sensor suite with sensors that that measure user generated forces could allow this system to identify and reject external disturbances while simultaneously shadowing user motion.
%
% We emphasize that user intent is critical, and current sensing challenges must be addressed in order to develop dynamic exoskeletons capable of serving in combat scenarios.  As a stop-gap, it may be possible to combine low information density EMG (thresholded on/off muscle activation signals) and human force/torque or angular measurements to filter user intent.  Such an approach may prove useful in a hybrid strategy, filtering external disturbance forces during sensitivity amplification control.  
% As a longer term solution, we highlight research in the area of MEMs and nano-fabrication, which has lead to the development of new, flexible ``robotic skin'' sensing arrays.  Integrating these sensing arrays into operator clothing (or attaching directly to the exoskeleton), it may be possible to directly measure human interaction forces (i.e., estimating pressure / force and exoskeleton contact locations), which was not feasible in the recent past.  Dense sensing arrays add redundancy and can improve measurement reliability. We recommend further study and the development of more advanced sensory systems to address the challenges inherent in dynamic exoskeleton control.
%
For instance, it may be possible to use low information density EMG (thresholded on/off muscle activation signals) with robot force/torque and angular measurements to filter disturbances from operator-robot interaction forces.  As a longer term solution, we highlight research in the area of MEMs, nano-fabrication, soft sensors, etc., which has lead to the development of new, flexible ``robotic skin'' sensing arrays.  Integrating these sensing arrays into operator clothing (or attaching directly to the exoskeleton), it may be possible to directly measure human interaction forces (i.e., estimating pressure / force and exoskeleton contact locations), which was not feasible in the recent past.  Dense sensing arrays add redundancy and can improve measurement reliability and, thereby, feedback control performance.

We also highly stress the importance of high-bandwidth control loops.  It is important that an exoskeleton's control bandwidth stays well above (typically $2-3\times$) the human control bandwidth ($2-5\,$Hz), to track dynamic motion.  Several of the reviewed exoskeletons controllers are close to lower limit of this human bandwidth threshold under no-load conditions, and so can interfere in dynamic tasks.  Note that these bandwidth issues are primarily a function of the exoskeleton hardware, and thus special care/analysis needs to be given/conducted to make sure that the hardware is capable of physically producing required behaviors.

Lastly, the importance of producing hardware that is physically capable of achieving dynamic performance is fundamentally linked to how mass is distributed throughout the system.  In general, it is desirable to minimize mass, particularly in distal regions. The design objective is often achieved by focusing actuation and leaving certain degrees of freedom passive (leading to an underactued system).  Underactuted control is a complicated problem that is often ignored in the sense that controllers assume the operator will provide necessary forces and torques for the un-actuated degrees of freedom.  Though it is possible to deal with the underactuated control problem directly (e.g., optimization-based controllers), there is a balance between feedback and feedforward control.  
% The better one is, the less reliant the system is on the other.  
By developing better sensors and feedback controllers that directly measure and shadow operator intent, exoskeleton controllers will become more robust and less sensitive to errors due to underactuated system assumptions and dynamic model inaccuracies.

%
%, or can be partially compensated using passive energy storage and dissipation elements.  
% we recommend a combination of biomechanically-inspired passive elements and optimization-based control methods, as these should ultimately produce the best strategy for dealing with underactuation in highly-dynamic exoskeleton systems.

@@

% Finally, we note that 
% Many of the common model-based control methods in this survey employ feedback linearization and inverse dynamics strategies for dynamic exoskeleton control.  
% These methods (incorrectly) assume the exoskeleton is fully actuated and require users to provide certain torques at uncontrolled joints.  There are more advanced control schemes that extend to underactuated systems and may apply here (e.g., optimization-based controllers or virtual model control).  However, we note that there is a balance between feedback and feedforward control.  The better one is, the less reliant the system is on the other.
%
% Improved feedback and direct sensing of user intent / human interaction forces therefore also offers to reduce sensitivity to high accuracy models and the fully actuated assumptions in inverse dynamics control.
